Sparse Representation Approaches for First Stage Retrieval: A Review

### 2022-01-03 src/article.tex
Update SNRM and SparTerm score function info.

    Therefore, a model is needed in which the representation sparsity is a function
    of the input length.
    Architecture based on Ngram representation learning.
    For each continuous n words in a given document or query, sparse representations are trained,
    which are then aggregated via the average pooling.
    Thus, the document similarity is represented as follows:
    \begin{equation}
        \phi_D(q)=\frac{1}{|d| - n + 1} \sum_{i=1}^{|d| - n + 1} \phi_{ngram}(w_i, w_{i+1}, \cdots, w_{i+n-1})
    \end{equation}
    where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\phi_{ngram}$ - learns a sparse representation
    for the given ngram.
    The same function is used for query representation.
    Ngram representation function is a fully-connected feed-forward network that uses a sliding
    window to read the input words and encodes their ngrams.
    Then embeddings are collected for each term \cite{SNRM}.

    \paragraph{Loss Function}
    The loss function for the $i^{th}$ training instance has the following form:
    \begin{equation}
        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi_Q(q_i)||\phi_D(d_{i1})||\phi_D(d_{i2}))
    \end{equation}
    where $\mathcal{L}$ - hinge loss \cite{IRNLP}, $q_i$ - query, $d_{i1}$ and $d_{i2}$ - two
    document candidates, $y_i \in \{-1, 1\}$ - label that shows the relevance between the document
    and the query, $\lambda$ - the hyper-parameter to control the sparsity of the learned
    representations, $L_1$ is norm (i.e.
    \begin{math}
        L_1(\vec{\upsilon}) = \sum_{i=1}^{|\vec{\upsilon}|}|\vec{\upsilon}_i|
    \end{math}
    ) and $||$ is concatenation sign.\\\par

@@

    \paragraph{Training the gating controller.}
    Consider a passage-query/summary parallel corpus \emph{C}, which contains passage \emph{p}, the
    corresponding target text \emph{t}, and the binary bag-of-words vector of \emph{t} (\emph{T}).
    In order to maximize probability values of all terms in vocabulary, binary cross-entropy loss
    is used, which has the following form:
    \begin{equation}
        L_{exp}=-\lambda_1\sum_{j \in \{m|T_m=0\}}log(1-G_j)-\lambda_2\sum_{k \in \{m|T_m=1\}}logG_k
    \end{equation}
    where $\lambda_1$ and $\lambda_2$ are tunable hyper-parameters, first is for terms expected not to be
    expanded, second one is for terms that appear in the target text, and \emph{G} is is the dense
    gating probability distribution for \emph{p} \cite{SparTerm}.\\\par
    Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words
    representation to a sparse term importance distribution in the whole vocabulary.
    This model also has limitation: since the gating mechanism is learned in advance, this prevents
    the model from learn the optimal sparsification strategy for the ranking task.

>>>>

    Therefore, a model is needed in which the representation sparsity is a function
    of the input length.
    Architecture based on Ngram representation learning.
    For each continuous n words in a given document or query, sparse representations are trained,
    which are then aggregated via the average pooling.
    Thus, the document similarity is represented as follows:
    \begin{equation}
        \psi(d)=
        \frac{1}{|d| - n + 1}
        \sum_{i=1}^{|d| - n + 1}
        \psi_{ngram}(w_i, w_{i+1}, \cdots, w_{i+n-1})
    \end{equation}
    where $w_1, w_2, \cdots, w_{|d|}$ - terms in document, $\psi_{ngram}$ - learns a sparse representation
    for the given ngram.

    The same function is used for query representation.

    The matching function for score is defined as the dot product of the document and query
    representations.

    Ngram representation function is a fully-connected feed-forward network that uses a sliding
    window to read the input words and encodes their ngrams.
    Then embeddings are collected for each term \cite{SNRM}.

    \paragraph{Loss Function}
    The loss function for the $i^{th}$ training instance has the following form:
    \begin{equation}
        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi(q_i)||\psi(d_{i1})||\psi(d_{i2}))
    \end{equation}
    where $\mathcal{L}$ - hinge loss \cite{IRNLP}, $q_i$ - query, $d_{i1}$ and $d_{i2}$ - two
    document candidates, $y_i \in \{-1, 1\}$ - label that shows the relevance between the document
    and the query, $\lambda$ - the hyper-parameter to control the sparsity of the learned
    representations, $L_1$ is norm (i.e.
    \begin{math}
        L_1(\vec{\upsilon}) = \sum_{i=1}^{|\vec{\upsilon}|}|\vec{\upsilon}_i|
    \end{math}
    ) and $||$ is concatenation sign.\\\par

@@

    \paragraph{Training the gating controller.}
    Consider a passage-query/summary parallel corpus \emph{C}, which contains passage \emph{p}, the
    corresponding target text \emph{t}, and the binary bag-of-words vector of \emph{t} (\emph{T}).
    In order to maximize probability values of all terms in vocabulary, binary cross-entropy loss
    is used, which has the following form:
    \begin{equation}
        L_{exp}=-\lambda_1\sum_{j \in \{m|T_m=0\}}log(1-G_j)-\lambda_2\sum_{k \in \{m|T_m=1\}}logG_k
    \end{equation}
    where $\lambda_1$ and $\lambda_2$ are tunable hyper-parameters, first is for terms expected not to be
    expanded, second one is for terms that appear in the target text, and \emph{G} is is the dense
    gating probability distribution for \emph{p} \cite{SparTerm}.

    \paragraph{Joint training.}
    The importance predictor and gating controller are trained jointly because the supervisory
    ranking signal can also be leveraged to guide the training of the gating controller:
    \begin{equation}
        L = L_{rank} + L_{exp}
    \end{equation}\\

    Thus, SparTerm provides a function that is used to map the frequency-based and bag-of-words
    representation to a sparse term importance distribution in the whole vocabulary.
    This model also has limitation: since the gating mechanism is learned in advance, this prevents
    the model from learn the optimal sparsification strategy for the ranking task.
